---
title: "Assignment9_SocialMedia"
author: "Akshita"
date: "2024-04-18"
output: html_document
---

```{r}
social_media <- read.csv("C:/Users/HP/Downloads/Social_media_original.csv")
str(social_media)
social_media_cleaned <- social_media[,-1]

#changing column names
change_cols_index <- c(2,4,6,8,10,12,14,16,17,18,19,20,21,22,23,24)
change_cols_name <- c("Instagram_Time", "Linkedin_Time", "Snapchat_Time", "Twitter_Time", "Whatsapp_Time", "Youtube_Time", "OTT_Time", "Reddit_Time", "Application Type", "Interview_call_received", "Networking", "Learning", "Mood_Productivity", "Morning_tireness", "Sleep_trouble", "Weekly_Feelings")
colnames(social_media_cleaned)[change_cols_index] <- change_cols_name


# Convert "NA", "N/A", "n/a", "na", "N.A", "n.a" to 0
social_media_cleaned[social_media_cleaned == "NA" | social_media_cleaned == "N/A" | social_media_cleaned == "na" | social_media_cleaned == "n/a" | social_media_cleaned == "N.A" | social_media_cleaned == "n.a" | social_media_cleaned == "0" | social_media_cleaned == ""] <- NA
social_media_cleaned[is.na(social_media_cleaned)] <- 0

# Define a function to convert time strings to decimal hours
convert_to_decimal_hours <- function(time_string) {
# Check if NA values are present
if (any(is.na(time_string))) {
         return(rep(NA, length(time_string)))  # Return NA for NA values
     }
     
# Define a function to convert HH:MM format to decimal hours
     hhmm_to_decimal <- function(hhmm) {
         parts <- as.numeric(strsplit(hhmm, ":")[[1]])  # Split into hours and minutes
         hours <- parts[1]
         minutes <- ifelse(length(parts) > 1, parts[2], 0)  # Handle missing minutes
         total_hours <- hours + minutes / 60
         return(total_hours)
     }
     
# Convert time strings to decimal hours
decimal_hours <- sapply(time_string, function(x) {
         if (grepl("^\\d+:\\d+$", x)) {
             return(hhmm_to_decimal(x))  # Convert HH:MM format
         } else if (grepl("^\\d+\\.\\d+$", x)) {
             return(as.numeric(x))  # Convert decimal format
         } else if (grepl("^\\d+$", x)) {
             return(as.numeric(x))  # Convert whole numbers
         } else {
             return(NA)  # Return NA for other cases
         }
     })
     
     return(decimal_hours)
}

time_columns <- c("Instagram_Time", "Linkedin_Time", "Snapchat_Time", "Twitter_Time", "Whatsapp_Time", "Youtube_Time", "OTT_Time", "Reddit_Time") 
# Apply the conversion function to all time columns
social_media_cleaned[time_columns] <- lapply(social_media_cleaned[time_columns], convert_to_decimal_hours)

#Treating NA in whatsapp column
mean_value <- mean(social_media_cleaned$Whatsapp_Time, na.rm = TRUE)
social_media_cleaned$Whatsapp_Time[is.na(social_media_cleaned$Whatsapp_Time)] <- mean_value

# Verify the result
str(social_media_cleaned)

# Convert 'No' and 'Yes' to 0 and 1 
social_media_cleaned$Mood_Productivity <- ifelse(social_media_cleaned$Mood_Productivity == "No", 0, 1)

social_media_cleaned$Morning_tireness <- ifelse(social_media_cleaned$Morning_tireness == "No", 0, 1)

social_media_cleaned$Sleep_trouble <- ifelse(social_media_cleaned$Sleep_trouble == "No", 0, 1)

# Select the relevant columns
time_columns <- social_media_cleaned[, grepl("_Time", colnames(social_media_cleaned))]
other_columns <- social_media_cleaned[c("Weekly_Feelings")]
 
# Combine the columns
combined_data <- cbind(time_columns, other_columns)
# Compute the correlation matrix
correlation_matrix <- cor(combined_data)
# Print the correlation matrix
print(correlation_matrix)
```

```{r}
#1. Model development
# Create a binary outcome variable
combined_data$Feelings_Category <- ifelse(combined_data$Weekly_Feelings > 3, "High", "Low")

# Convert Feelings_Category to factor
combined_data$Feelings_Category <- factor(combined_data$Feelings_Category, levels = c("Low", "High"))

# Check the structure of the dataset
str(combined_data)

# Drop Weekly_Feelings variable using subset function
combined_data <- subset(combined_data, select = -Weekly_Feelings)

# Fit logistic regression model
logistic_model <- glm(Feelings_Category ~ ., data = combined_data, family = binomial)

# Summary of the model
summary(logistic_model)

#2. Model Acceptance & 3. Residual Analysis:
# Check p-values
p_values <- summary(logistic_model)$coefficients[, 4]
print(p_values)

# Check AIC and BIC
AIC_value <- AIC(logistic_model)
print(AIC_value)
BIC_value <- BIC(logistic_model)
print(BIC_value)

# Compare null deviance and residual deviance
null_deviance <- summary(logistic_model)$null.deviance
residual_deviance <- summary(logistic_model)$deviance
deviance_reduction <- null_deviance - residual_deviance
print(paste("Null Deviance:", null_deviance))
print(paste("Residual Deviance:", residual_deviance))
print(paste("Deviance Reduction:", deviance_reduction))

# Assumptions Checking (Diagnostic Plots)
par(mfrow = c(2, 2))  # Set up a 2x2 plot layout
plot(logistic_model)

# Plot residuals
plot(resid(logistic_model), ylab = "Residuals")

#4. Prediction
# Predicting probabilities
predicted_probs <- predict(logistic_model, type = "response")

#5, Model Accuracy
# Create a confusion matrix
predicted_classes <- ifelse(predicted_probs > 0.5, "High", "Low")
conf_matrix <- table(predicted_classes, combined_data$Feelings_Category)

# Calculate accuracy
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)
print(accuracy)

# Load required library
library(pROC)

# Calculate ROC curve
roc_curve <- roc(combined_data$Feelings_Category, predicted_probs)

# Plot ROC curve
plot(roc_curve, main = "ROC Curve", col = "blue")

# Add diagonal reference line
abline(a = 0, b = 1, lty = 2, col = "red")

# Add AUC to the plot
auc_value <- auc(roc_curve)
text(0.8, 0.2, paste("AUC =", round(auc_value, 2)), col = "black")

```
1. A one-unit increase in "Instagram_Time" is associated with a 0.14075 increase in the log odds of the feelings category, holding other variables constant.

In this output, "Linkedin_Time" has a p-value of 0.0909, indicating it is marginally significant at the 0.1 significance level.

A significant reduction in deviance indicates that the predictors improve the model's fit.
The AIC value provides a measure of the model's goodness of fit, with lower values indicating a better fit.

2. p-values:
Most predictor variables have p-values > 0.05, indicating they are not statistically significant.
"Linkedin_Time" has a p-value < 0.1, suggesting potential significance.
Other variables may require further investigation or elimination.

AIC and BIC:
AIC: 35.93, indicating reasonable fit and parsimony.
BIC: 45.33, with lower values preferred.

Deviance:
Reduction in deviance from null to residual: 9.98, indicating substantial explained variation.
```{r}
#improved accuracy

# Load required library
library(glmnet)

# Prepare the data
X <- as.matrix(combined_data[, -9]) # Independent variables
Y <- as.factor(combined_data$Feelings_Category) # Dependent variable

# Perform cross-validation to tune lambda (regularization parameter)
cv_model <- cv.glmnet(X, Y, family = "binomial", alpha = 0)

# Get the best lambda
best_lambda <- cv_model$lambda.min

# Fit the final model with the best lambda
final_model <- glmnet(X, Y, family = "binomial", alpha = 0)
best_model <- glmnet(X, Y, family = "binomial", alpha = 0, lambda = best_lambda)

# Cross-validation accuracy
cv_accuracy <- 1 - cv_model$cvm[length(cv_model$cvm)] # Extract the accuracy from the last fold

# Predicting probabilities
predicted_probs <- predict(best_model, newx = X, type = "response")

# Model Accuracy
# Create a confusion matrix
predicted_classes <- ifelse(predicted_probs > 0.5, "High", "Low")
conf_matrix <- table(predicted_classes, Y)

# Calculate accuracy
accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)

# Print accuracy
print(accuracy)

# Previous accuracy
previous_accuracy <- 0.1904762

# New accuracy
new_accuracy <- accuracy

# Create a bar plot
barplot(c(previous_accuracy, new_accuracy), 
        names.arg = c("Previous", "New"), 
        ylim = c(0, 1), 
        col = c("red", "blue"),
        main = "Comparison of Model Accuracies",
        ylab = "Accuracy")

# Add text labels
text(1, previous_accuracy + 0.05, round(previous_accuracy, 2), pos = 3, col = "red")
text(2, new_accuracy + 0.05, round(new_accuracy, 2), pos = 3, col = "blue")

```
